---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS litters data

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

##  `select`

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming column ...

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers

?select_helpers

select columns that start with "gd"
```{r}
select(litters_df, starts_with("gd"))
```

move litter_number to the beginning and keep anything else the same
```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df, litter_number)
```

## `filter`

choose rows

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

== is a test to equality
```{r}
filter(litters_df, gd_of_birth == 20)
```

not equal 
```{r}
filter(litters_df, !(gd_of_birth == 20))
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

choose in either group
```{r}
filter(litters_df, group %in% c("Con7","Mod8"))
```

## `mutate`

change existing variable and create new variable

create a new variable wt_gain
change group to lowercase
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

