---
title: "Exploratory Data Analysis"
output: github_document
---

```{r}
library("dplyr")
```

## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplots.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_continuous = scale_fill_viridis_d
```


## load the dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2023-12-31")  %>% 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

month = lubridate::floor_date(date, unit = "month")) : round date to the month


## 'group_by'

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup(month)
```

## counting things

count months/ name observations

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

n(): count the number

we can use 'count()'

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

**NEVER** use base R's `table`

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

The result is not a data frame and not helpful

other helpful counters
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

n_distince(date): count the number of distinct observations.e.g. in 2021-01-01, how many distinct days in this month (31).


## A digression on 2*2 table

```{r}
weather_df %>% 
  filter(name !="Molokai_HI") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

case_when: given this condition use this output level

```{r}
weather_df %>% 
  filter(name !="Molokai_HI") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)
```
