---
title: "Visualization"
subtitle: "Part 1"
output: github_document
---


install.packages("ggridges")

install.packages(c('crul', 'XML', 'gridExtra', 'isdparser', 'geonames', 'hoardr'))
install.packages("~/Downloads/rnoaa_1.4.0.tar.gz", repos = NULL, type = "source")


```{r setup}
library(tidyverse)
library(ggridges)
```


## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") %>% 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10)  %>% 
  select(name, id, everything())
```

-metep_pull_monitors: pull data from open resources


## Scatterplots

Create my first scatterplot ever

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

-aes: stands for "aesthetics" and is used to specify the mapping of variables in your data to visual properties (aesthetics) of the plot. This includes properties like x and y coordinates, colors, shapes, sizes, etc.

-geom_point():adds a geometric layer to your plot. In this case, geom_point creates a scatter plot .

New approach, same plot.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit a plot object.

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot

Start with the same one and make it fancy!
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

-color points according the name.
-geom_smooth(se = FALSE): add smooth curve.
-se = FALSE: Disables the display of the confidence interval. By default, geom_smooth shows a 95% confidence interval around the smooth curve.

What about the aes placement ?
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

- color only applied to the points and not include geom_smooth.

Let's facet some things!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2, size = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

-ggplot(aes(x = tmin, alpha = tmin, y = tmax, color = name)): points transparency will change based on tmin
-geom_point(alpha = .2, size = .3): make points 20% transparency, size to .3
-facet_grid(. ~ name): don't create rows, but create columns that describe names: three panel plots
-facet_grid(name ~ .): create rows (three panel plots according to names), but not columns


Let's combine some elements and try a new plot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```
-aes(size = prcp): point size shows the prcp(precipitation).



## Some small notes

How many geoms have to exist?

You can have whatever geoms you want.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

-geom_hex: give us 2D plot(hexagrams). When sample size is large, this helps is to understand the distribution of the data. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
```

-geom_bin2d: You can also use geom_bin2d, which is in square.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = .3)
```

geom_density2d: like a contour plot


## Univariate plots

Histograms are really great.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

-position = "dodge":it separates the bars for different groups so that they do not overlap but are placed next to each other. 
-fill = name: fill the color in the bar (color = name just color the outside of the bar)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

-facet_grid(. ~ name): create three panel plots separately

Let's try a new geometry!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3, adjust = .5)
```

geom_density: density curve
adjust = .5: adjusts the bandwidth of the density estimate.Give you more information about the change of the trend(more bumps).

What about box plots?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots :-)

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary(fun = "median")
```

-geom_violin: violin plot, might be more informative (more distribution shape) than box plot
-stat_summary(fun = "median"): show the median


Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```
geom_density_ridges: Ridge plots are used to visualize the distribution of a numeric variable for several groups. Each group’s distribution is displayed as a density plot, and the plots are stacked vertically.


