---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

install.packages("patchwork")
install.packages("viridis")

```{r}
library(tidyverse)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2023-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```


## Remember this plot ..?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

## Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  )
```

## Scale

Start with the same plot; x and y scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15 C")
  ) +
  scale_y_continuous(
    position = "right"
  )
```

-scale_x_continuous: change x axis labels
-scale_y_continuous: change y axis labels
-trans = "sqrt":square root transformation, you can also do trans = "log"
- position = "right": put y labels on the right side

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(150,260))
```

-scale_color_hue:map the color aesthetic to hues in the HCL (hue-chroma-luminance) color space. 
-h = c(10, 260): Specifies the range of hues to be used. The hue is an angle in the color wheel, with values ranging from 0 to 360. 


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE
 )
```

viridis::scale_color_viridis: give different color to locations
discrete = TRUE: viridis default thinks scales are continue so here need to tell R that scale is discrete.


## Themes

Shift the legend

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE) +
  theme(legend.position = "bottom")
```

legend.position = "bottom": shift the legend from side to the bottom

Change the overall theme
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2021 to 2023"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

theme_minimal: change the background into white, and without the box. 
theme_bw():change the background into white
theme_classic():white background and no gridlines
ggthemes::theme_economist: there are many other theme in ggthemes


## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplots.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_continuous = scale_fill_viridis_d
```

eval = FALSE: code within the block will not be executed.
If I have this in my very first code chunk in the markdown document, this will load tidyverse and update the theme everywhere.
from options ... scale_fill_viridis_d: make sure plots use the virtus color palette
Put this in your Rmarkdown file and unit the theme of plots.


## Data args in 'geom'
```{r}
central_park =
  weather_df %>%
  filter(name == "CentralPark_NY")

moloka =
  weather_df %>%
  filter(name == "Molokai_HI")

ggplot(data = moloka, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

Apply different geom to different datasets.
If I delete the data = central_park, geom_line will apply to moloka.


## 'patchwork'

remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

what happens when you want multipanel plots but can't facet?

```{r}
tmax_tmin_p =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp >0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5)

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

tmax_tmin_p + prcp_dens_p + tmax_date_p
tmax_tmin_p / (prcp_dens_p + tmax_date_p)
(tmax_tmin_p + prcp_dens_p) / tmax_date_p

```

tmax_tmin_p + prcp_dens_p + tmax_date_p: put three plots together
tmax_tmin_p / (prcp_dens_p + tmax_date_p): change the position of plots


## Data manipulation

Control your factors

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Molokai_HI"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
    
  
```
name = factor(name): change name from character to factor(categorical data).
name = forcats::fct_relevel(name, c("Molokai_HI")): number the Molokai as the first and others stay the same.

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```
 
 
## Revisit the pups

Data from the FAS study.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, '1' = "male", '2' = "female"))

litter_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>%  
  select(dose, day_of_tx, starts_with("pd_")) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day"
  ) %>%  
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median))%>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```



